Incentivizing Exploration by Heterogeneous Users

### 2018-02-15 algorithm.tex
halfway through editing payment proof

The algorithm achieving the claimed bounds of
Theorem~\ref{thm:main-overview} is fairly simple.
It mostly allows agents to exploit, but when an arm is \pfdelete{estimated to
be extremely} \pfedit{sufficiently} unlikely to be pulled\pfdelete{by any agents},
the algorithm incentivizes this arm with a payment high enough
to guarantee the next agent pulls it.
This way the algorithm ensures each arm is pulled often enough.

@@

\begin{itemize}
\item $i$ has been pulled at most%
\footnote{in fact: exactly, since the algorithm entered phase $s$}
$s-1$ times up to time $t$, i.e., $\NumPull{t}{s} \leq s-1$.
\item Based on the current estimates \ArmEV{t}{i'} of all arms'
attribute vectors, the probability of pulling arm $i$ is less than $1/\log(s)$. \bccomment{Do we need to mention what if $\frac{1}{\log(s)}>1$? }
Let $\PullProb{t}{i} = \Prob[\AgV \sim \AgentDist]{\AgV \cdot \ArmEV{t}{i} > \AgV
  \cdot \ArmEV{t}{i'} \mbox{ for all } i' \neq i}$
be the probability that arm $i$ will be pulled
by the next (random) agent based on the current estimates. 
Then, the second criterion for payment-eligibility is that
$\PullProb{t}{i} < 1/\log(s)$.
\pfedit{$1/\log(1)$ is treated as $\infty$, and if $1/\log(s) \ge 1$ then this condition is automatically met.}
\end{itemize}

@@

In this section, we prove Algorithm~\ref{alg:basic-incentivizing} achieves $O(\ARMNUM^2+\TieDensity(\log(T))^2)$ cumulative regret with $O(\ARMNUM^2)$ payment budget.  This is stated in the following pair of theorems, which together constitute our main results.

@@

In either case, the expected number of time steps until
$|S|$ decreases is at most $s$,
so the expected number of steps until $S$ is empty is at most
$\ARMNUM \cdot \log(s)$.
\dkcomment{If we wanted, this bound could probably be improved using a
  coupon collector argument. But it would look uglier.}
\end{proof}

@@

\begin{lemma}[Corollary 1 of \cite{zhao2016adaptive}]
\label{lem:ACI-inequality}
Let $X_i$ be zero-mean $1/2$-subgaussian random variables,
and $\SET{S_n = \sum_{i=1}^n X_i, n \geq 1}$ the corresponding random walk.
Let $J$ be any stopping time with respect to $\SET{X_1, X_2, \ldots}$.
(We allow $J$ to take the value $\infty$,
defining $\Prob{J = \infty} = 1 - \lim_{n \to \infty} \Prob{J \leq n}$.)
Define \dkcomment{I'm making the dependence of $b$ on $n$ explicit,
  since that's how we're using it.}
\begin{align*}
g(n) & = \sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + n \cdot b(n)}.
\end{align*}
Then, 
\begin{align*}
\Prob{J < \infty \mbox{ and } S_J \geq g(J)} \leq 12 \e^{-1.8 b(n)}.
\end{align*}
\end{lemma}

@@

Then, we use the following technical Lemma~\ref{lem:n0-inequality}
(whose proof is given in the appendix, and whose precondition we verify
momentarily) to bound the first term,
giving us an overall upper bound of
$|S_J| \leq \frac{J x}{2 \sigma}$.

@@

It would be desirable to simply identify a constant upper bound on the
payment made each time.
Unfortunately, while the agent types are drawn from a bounded support,
the noise in arm locations is not;
hence, with small probability, arm locations may be grossly
misestimated, resulting in high incentive payments.
As a result, the actual analysis of the total payment is significantly more
intricate, and hence,
the proof of Theorem~\ref{rst:budget} is given in the appendix.

@@

The first line is independent of $T$, and we now choose
$g^2(s) = \frac{288\log(T) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}$
with the goal of minimizing the second line.
With this choice and same definition of $S_{s}$ as in the proof of Lemma~\ref{lem:round-prob}, 
\begin{align*}
&\Prob{\exists i,j, s.t. |\ArmE{t}{i}{j} - \Arm{i}{j}|\geq \frac{g(s)}{2\Diam d}} \\
\leq &\sum_{i,j}\Prob{|S_s|\geq 3\sqrt{\frac{2}{1.8}s\log(T)}} \\
\leq &\sum_{i,j}\Prob{|S_s|\geq \sqrt{0.6s\log(\log_{1.1}(s)+1)+\frac{2}{1.8}\log(T)s}} \leq 24\ARMNUM d \cdot \exp(-2\log(T)),
\end{align*}
where the second to last inequality is because the following technical inequality, whose proof we include in the appendix.

>>>>

The algorithm achieving the claimed bounds of
Theorem~\ref{thm:main-overview} is fairly simple.
It mostly allows agents to exploit, but when an arm is sufficiently
unlikely to be pulled, 
the algorithm incentivizes this arm with a payment high enough
to guarantee that the next agent pulls it.
This way the algorithm ensures each arm is pulled often enough.

@@

\begin{itemize}
\item $i$ has been pulled at most%
\footnote{in fact: exactly, since the algorithm entered phase $s$}
$s-1$ times up to time $t$, i.e., $\NumPull{t}{s} \leq s-1$.
\item Based on the current estimates \ArmEV{t}{i'} of all arms'
attribute vectors, the probability of pulling arm $i$ is less than
$1/\log(s)$.
\bccomment{Do we need to mention what if $\frac{1}{\log(s)}>1$?}
\dkcomment{I'm not sure why. It's not like the inequality is not
  satisfied.}
Let $\PullProb{t}{i} = \Prob[\AgV \sim \AgentDist]{\AgV \cdot \ArmEV{t}{i} > \AgV
  \cdot \ArmEV{t}{i'} \mbox{ for all } i' \neq i}$
be the probability that arm $i$ will be pulled
by the next (random) agent based on the current estimates. 
Then, the second criterion for payment-eligibility is that
$\PullProb{t}{i} < 1/\log(s)$.
\pfedit{$1/\log(1)$ is treated as $\infty$, and if $1/\log(s) \ge 1$ then this condition is automatically met.}
\dkcomment{Do we need this?}
\end{itemize}

@@

In this section, we prove that Algorithm~\ref{alg:basic-incentivizing} achieves $O(\ARMNUM^2+\TieDensity(\log(T))^2)$ cumulative regret with $O(\ARMNUM^2)$ payment budget.  This is stated in the following pair of theorems, which together constitute our main results.

@@

In either case, the expected number of time steps until
$|S|$ decreases is at most $\log s$,
so the expected number of steps until $S$ is empty is at most
$\ARMNUM \cdot \log(s)$.
\end{proof}

@@

\begin{lemma}[Corollary 1 of \cite{zhao2016adaptive}]
\label{lem:ACI-inequality}
Let $X_i$ be zero-mean $1/2$-subgaussian random variables,
and $\SET{S_n = \sum_{i=1}^n X_i, n \geq 1}$ the corresponding random walk.
Let $J$ be any stopping time with respect to $\SET{X_1, X_2, \ldots}$.
(We allow $J$ to take the value $\infty$,
defining $\Prob{J = \infty} = 1 - \lim_{n \to \infty} \Prob{J \leq n}$.)
Define \dkcomment{I'm making the dependence of $b$ on $n$ explicit,
  since that's how we're using it. Is it true with $b$ depending on
  $n$?
If not, we need to revisit what we're doing.}
\begin{align*}
g(n) & = \sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + n \cdot b(n)}.
\end{align*}
Then, 
\begin{align*}
\Prob{J < \infty \mbox{ and } S_J \geq g(J)} \leq 12 \e^{-1.8 b(n)}.
\end{align*}
\end{lemma}

@@

Then, we use the following technical Lemma~\ref{lem:n0-inequality}
(whose proof is given in Appendix~\ref{n0-inequality-proof}, and whose
precondition we verify momentarily) to bound the first term,
giving us an overall upper bound of
$|S_J| \leq \frac{J x}{2 \sigma}$.

@@

It would be desirable to simply identify a constant upper bound on the
payment made each time.
Unfortunately, while the agent types are drawn from a bounded support,
the noise in arm locations is not;
hence, with small probability, arm locations may be grossly
misestimated, resulting in high incentive payments.
As a result, the actual analysis of the total payment is significantly more
intricate, and hence,
the proof of Theorem~\ref{rst:budget} is given in
Appendix~\ref{sec:payment-proof}.

@@

The first line is independent of $T$, and we now choose
$g^2(s) = \frac{288\log(T) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}$
with the goal of minimizing the second line.
With this choice and same definition of $S_{s}$ as in the proof of Lemma~\ref{lem:round-prob}, 
\begin{align*}
&\Prob{\exists i,j, s.t. |\ArmE{t}{i}{j} - \Arm{i}{j}|\geq \frac{g(s)}{2\Diam d}} \\
\leq &\sum_{i,j}\Prob{|S_s|\geq 3\sqrt{\frac{2}{1.8}s\log(T)}} \\
\leq &\sum_{i,j}\Prob{|S_s|\geq \sqrt{0.6s\log(\log_{1.1}(s)+1)+\frac{2}{1.8}\log(T)s}} \leq 24\ARMNUM d \cdot \exp(-2\log(T)),
\end{align*}
where the second to last inequality is because the following technical inequality, whose proof we include in Appendix~\ref{sec:n1-inequality-proof}.
